The Correspondence Analysis of Two Mode Networks Revisited

### 2025-02-20 analysis-R1.qmd
plots!

```{r Faust-Style Direct CA}
   ca.res <- svd(diag(1/sqrt(rowSums(A))) %*% A %*% diag(1/sqrt(colSums(A))))
   d1.r <- ca.res$d[2] * ca.res$u[, 2] * sqrt(N/rowSums(A))
   d1.c <- ca.res$d[2] * ca.res$v[, 2] * sqrt(N/colSums(A))
   d1 <- c(d1.r, d1.c)
   d2.r <- ca.res$d[3] * ca.res$u[, 3] * sqrt(N/rowSums(A))
   d2.c <- ca.res$d[3] * ca.res$v[, 3] * sqrt(N/colSums(A))
   d2 <- c(d2.r, d2.c)
   d3.r <- ca.res$d[4] * ca.res$u[, 4] * sqrt(N/rowSums(A))
   d3.c <- ca.res$d[4] * ca.res$v[, 4] * sqrt(N/colSums(A))
   d3 <- c(d3.r, d3.c)
   d4.r <- ca.res$d[5] * ca.res$u[, 5] * sqrt(N/rowSums(A))
   d4.c <- ca.res$d[5] * ca.res$v[, 5] * sqrt(N/colSums(A))
   d4 <- c(d4.r, d4.c)
```

@@

```{Two Dual Projection Eigenvector Scores}
   eig.r <- eigen(A %*% t(A))$vectors
   eig.c <- eigen(t(A) %*% A)$vectors
   rownames(eig.r) <- rownames(A)
   rownames(eig.c) <- colnames(A)
```

@@

```{r CA versus Reflections Scatter}
   p.plot.dat <- data.frame(c = tm$ca.eigvec.r[, 1]*-1, r = a$p.s[, 26])
   p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
   p <- p + geom_point(size = 3, color = "blue") 
   p <- p + geom_text_repel(aes(label = rownames(p.plot.dat)), color = "red")
   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "26th Reflection (Standardized)")
   p <- p + theme(axis.text = element_text(size = 16),
                  axis.title = element_text(size = 16))
   png(file = here("Plots", "p-ca-ref-corr.png"), width=600, height=625)
   p
   dev.off()

   g.plot.dat <- data.frame(cbind(c = tm$ca.eigvec.c[, 1]*-1, r = a$g.s[, 26]))
   p <- ggplot(data = g.plot.dat, aes(x = c, y = r))
   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
   p <- p + geom_point(size = 3, color = "red") 
   p <- p + geom_text_repel(aes(label = rownames(g.plot.dat)), 
                            color = "blue", size = 6)
   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "26th Reflection (Standardized)")
   p <- p + theme(axis.text = element_text(size = 16),
                  axis.title = element_text(size = 16))
   png(file = here("Plots", "g-ca-ref-corr.png"), width=600, height=625)
   p
   dev.off()
```

@@

```{r Row Stochastic Matrices for CA Projections Plot}
   D.p <- diag(1/rowSums(A)) #inverse of degree matrix of persons
   P.pg <- D.p %*% A
   rownames(P.pg) <- rownames(A) 
   D.g <- diag(1/colSums(A)) #inverse of degree matrix of groups
   P.gp <- D.g %*% t(A)
   rownames(P.gp) <- colnames(A)
   P.pp <- P.pg %*% P.gp
   P.gg <- P.gp %*% P.pg
   rownames(P.pp) <- substr(colnames(P.pp), start = 1, stop = 3)
   rownames(P.gg) <- substr(colnames(P.gg), start = 1, stop = 3)
   P.p <- ggcorrplot(P.pp, lab = TRUE) +
      theme(legend.position = "none", 
            axis.text.x = element_text(hjust = -0.2)) + 
      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
      scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5))
   P.g <- ggcorrplot(P.gg, lab = TRUE) +
      theme(legend.position = "none", 
            axis.text.x = element_text(hjust = -0.2)) + 
      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) +
      scale_fill_gradient2(breaks = c(0, .5), limit = c(0, .5))
   png(file = here("Plots", "p-norm.png"), width=600, height=500)
   P.p
   dev.off() 
   png(file = here("Plots", "g-norm.png"), width=600, height=500)
   P.g
   dev.off() 
```

```{r CA Eigenvector Plot} 
      eigvec.dat.r <- data.frame(rank = rank(d1.r), 
                             value = as.numeric(d1.r),
                             lab = names(d1.r))
      eigvec.dat.c <- data.frame(rank = rank(d1.c), 
                             value = as.numeric(d1.c),
                             lab = names(d1.c))
      eigval.dat.r <- data.frame(k = 1:nrow(A)-1, value = as.numeric(CA.r$values)) %>% 
             mutate(value = value / max(value))
      eigval.dat.c <- data.frame(k = 1:ncol(A)-1, value = as.numeric(CA.c$values)) %>% 
             mutate(value = value / max(value))

@@

      # Eigenvector plot for the groups
      p <- ggscatter(eigvec.dat.c, y = "rank", x = "value", point = FALSE,
                 font.label =  10, label = "lab", repel = TRUE)
      p <- p + theme(axis.text.y = element_blank(),
                 axis.line.x = element_blank(),
                 axis.line.y = element_blank(),
                 axis.ticks = element_blank(),
                 legend.position = "none")
      p <- p + labs(y = "Rank", x = "First Axis") 
      b <- p + geom_vline(xintercept = 0, linewidth = 0.5, linetype = 2, color = "gray")

      # Eigenvalue plot for the people
      p <- ggscatter(eigval.dat.r, x = "k", y = "value", size = 2, color = "black")
      p <- p + labs(y = "", x = "k") 
      c <- p + theme(axis.text.x = element_blank(),
                    axis.text.y = element_text(size = 8),
                    axis.line.y = element_blank(),
                    axis.line.x = element_blank())

      # Eigenvalue plot for the groups
      p <- ggscatter(eigval.dat.c, x = "k", y = "value", size = 2, color = "black")
      p <- p + labs(y = "", x = "k") 
      d <- p + theme(axis.text.x = element_blank(),
                    axis.text.y = element_text(size = 8),
                    axis.line.y = element_blank(),
                    axis.line.x = element_blank())
      a <- a + inset_element(c, left = 0.7, right = 1, bottom = 0, top = 0.45)
      b <- b + inset_element(d, left = 0.7, right = 1, bottom = 0, top = 0.45)
      png(file = here("Plots", "p-ca-eigvec.png"), width=500, height=500)
      a
      dev.off() 
      png(file = here("Plots", "g-ca-eigvec.png"), width=500, height=500)
      b
      dev.off() 
```

```{r CA Similarity and Correspondence Plot}
      norm.sim.r <- as.matrix(S.r/max(S.r))
      norm.sim.c <- as.matrix(S.c/max(S.c))
      norm.sim.r <- norm.sim.r[order(d1.r), order(d1.r)]
      norm.sim.c <- norm.sim.c[order(d1.c), order(d1.c)]

@@

      # CA Correspondence Plot
      plot.dat <- data.frame(a = d3, 
                             b = d1, 
                             mode = c(rep(1, 18), rep(2, 14)))
      p <- ggplot(plot.dat, aes(y = a, x = b, color = factor(mode), label = rownames(plot.dat)))
      p <- p + geom_vline(xintercept = 0, color = "gray")
      p <- p + geom_hline(yintercept = 0, color = "gray")
      p <- p + geom_text_repel(size = 4, max.overlaps = 50)
      p <- p + theme_minimal() 
      p <- p + theme(legend.position = "none",
                     axis.text = element_text(size = 18),
                     axis.title = element_text(size = 20)
                     )
      p <- p + labs(x = "Second Dimension", y = "First Dimension")
      c <- p + scale_color_manual(values = c('red', 'blue'))
      png(file = here("Plots", "ca-corr-plot.png"), width=1200, height=900)
      (a + b) / c + 
         plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
         theme(plot.tag = element_text(size = 26))
      dev.off()
```

>>>>

```{r Faust-Style Direct CA}
   FCA <- svd(diag(1/sqrt(rowSums(A))) %*% A %*% diag(1/sqrt(colSums(A))))
   w.r <- sqrt(sum(A)/rowSums(A))
   w.c <- sqrt(sum(A)/colSums(A))
   FCA$d <- FCA$d[2:14]
   FCA$u <- FCA$u[, 2:14]
   FCA$v <- FCA$v[, 2:14]
   #FCA$u <- FCA$u * FCA$d * w.r
   #FCA$v <- FCA$v * FCA$d * w.c
   FCA.d <- rbind(FCA$u, FCA$v)
   rownames(FCA.d) <- c(rownames(A), colnames(A))
   round(FCA.d, 3)
```

@@

```{r Dual Projection Eigenvector Scores}
   eig.r <- eigen(A %*% t(A))$vectors
   eig.c <- eigen(t(A) %*% A)$vectors
   rownames(eig.r) <- rownames(A)
   rownames(eig.c) <- colnames(A)
```

@@

```{r CA versus Reflections Scatter}
   p.plot.dat <- data.frame(c = CA.r$vectors[, 2]*-1, 
                            r = a$p.s[, 26])
   p <- ggplot(data = p.plot.dat, aes(x = c, y = r))
   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
   p <- p + geom_point(size = 3, color = "blue") 
   p <- p + geom_text_repel(aes(label = rownames(p.plot.dat)), color = "red")
   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "26th Reflection (Standardized)")
   p <- p + theme(axis.text = element_text(size = 16),
                  axis.title = element_text(size = 16))
   png(file = here("Plots", "p-ca-ref-corr.png"), width=600, height=625)
   p
   dev.off()

   g.plot.dat <- data.frame(cbind(c = CA.c$vectors[, 2]*-1, 
                                  r = a$g.s[, 26]))
   p <- ggplot(data = g.plot.dat, aes(x = c, y = r))
   p <- p + geom_smooth(method = "lm", se=FALSE, formula = y ~ x, color = "gray")
   p <- p + geom_point(size = 3, color = "red") 
   p <- p + geom_text_repel(aes(label = rownames(g.plot.dat)), 
                            color = "blue", size = 6)
   p <- p + theme_minimal() + labs(x = "First CA Dimension", y = "26th Reflection (Standardized)")
   p <- p + theme(axis.text = element_text(size = 16),
                  axis.title = element_text(size = 16))
   png(file = here("Plots", "g-ca-ref-corr.png"), width=600, height=625)
   p
   dev.off()
```

@@

```{r Row Stochastic Matrices for CA Projections Plot}
   P.pg <- diag(1/rowSums(A)) %*% A
   rownames(P.pg) <- rownames(A) 
   P.gp <- diag(1/colSums(A)) %*% t(A)
   rownames(P.gp) <- colnames(A)
   P.pp <- P.pg %*% P.gp
   P.gg <- P.gp %*% P.pg
   rownames(P.pp) <- substr(colnames(P.pp), start = 1, stop = 3)
   rownames(P.gg) <- substr(colnames(P.gg), start = 1, stop = 3)
   P.p <- ggcorrplot(P.pp, lab = TRUE, lab_size = 5,
                     colors = c(rep("white", 3))) +
      theme(legend.position = "none", 
            axis.text.x = element_text(hjust = -0.2, size = 16),
            axis.text.y = element_text(size = 16)) + 
      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) 
   P.g <- ggcorrplot(P.gg, lab = TRUE, lab_size = 5,
                     colors = c(rep("white", 3))) +
      theme(legend.position = "none", 
            axis.text.x = element_text(hjust = -0.2, size = 16),
            axis.text.y = element_text(size = 16)) + 
      scale_x_discrete(position = "top") + scale_y_discrete(limits=rev) 
   png(file = here("Plots", "p-norm.png"), width=900, height=900)
   P.p
   dev.off() 
   png(file = here("Plots", "g-norm.png"), width=900, height=900)
   P.g
   dev.off() 
```

```{r CA Eigenvector Plot} 
      eigvec.dat.r <- data.frame(rank = rank(CA.r$vectors[,2]), 
                             value = as.numeric(CA.r$vectors[,2]),
                             lab = names(CA.r$vectors[,2]))
      eigvec.dat.c <- data.frame(rank = rank(CA.c$vectors[,2]), 
                             value = as.numeric(CA.c$vectors[,2]),
                             lab = names(CA.c$vectors[,2]))
      eigval.dat.c <- data.frame(k = 1:ncol(A)-1, 
                                 value = as.numeric(CA.c$values)) %>% 
             mutate(value = value / max(value))

@@

      # Eigenvector plot for the groups
      p <- ggscatter(eigvec.dat.c, y = "rank", x = "value", point = FALSE,
                 font.label =  10, label = "lab", repel = TRUE)
      p <- p + theme(axis.text.y = element_blank(),
                 axis.line.x = element_blank(),
                 axis.line.y = element_blank(),
                 axis.ticks = element_blank(),
                 plot.margin = margin(l = 100),
                 legend.position = "none")
      p <- p + labs(y = "Rank", x = "First Axis") 
      b <- p + geom_vline(xintercept = 0, linewidth = 0.5, linetype = 2, color = "gray")

      # Eigenvalue plot for the groups
      p <- ggscatter(eigval.dat.c, x = "k", y = "value", size = 2, color = "black")
      p <- p + labs(y = "", x = "k") 
      c <- p + theme(axis.text.x = element_blank(),
                    axis.text.y = element_text(size = 8),
                    axis.line.y = element_blank(),
                    axis.line.x = element_blank())
      b <- b + inset_element(c, left = 0.6, right = 1, bottom = 0, top = 0.45)
      png(file = here("Plots", "ca-eigvec.png"), width=1000, height=600)
      a + b + plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
         theme(plot.tag = element_text(size = 20))
      dev.off() 
```

```{r CA Similarity and Correspondence Plot}
      norm.sim.r <- as.matrix(P.pp/max(P.pp))
      norm.sim.c <- as.matrix(P.gg/max(P.gg))
      norm.sim.r <- norm.sim.r[order(CA.r$vectors[,2]), 
                               order(CA.r$vectors[,2])]
      norm.sim.c <- norm.sim.c[order(CA.c$vectors[,2]), 
                               order(CA.c$vectors[,2])]

@@

      # CA Correspondence Plot
      plot.dat <- data.frame(a = FCA.d[, 2], 
                             b = FCA.d[, 1], 
                             mode = c(rep(1, 18), rep(2, 14)))
      p <- ggplot(plot.dat, aes(y = a, x = b, color = factor(mode), label = rownames(plot.dat)))
      p <- p + geom_vline(xintercept = 0, color = "gray")
      p <- p + geom_hline(yintercept = 0, color = "gray")
      p <- p + geom_text_repel(size = 4, max.overlaps = 50)
      p <- p + theme_minimal() 
      p <- p + theme(legend.position = "none",
                     axis.text = element_text(size = 18),
                     axis.title = element_text(size = 20)
                     )
      p <- p + labs(x = "First Dimension", y = "Second Dimension")
      c <- p + scale_color_manual(values = c('red', 'blue'))

      plot.dat <- data.frame(a = FCA.d[, 3], 
                             b = FCA.d[, 1], 
                             mode = c(rep(1, 18), rep(2, 14)))
      p <- ggplot(plot.dat, aes(y = a, x = b, color = factor(mode), label = rownames(plot.dat)))
      p <- p + geom_vline(xintercept = 0, color = "gray")
      p <- p + geom_hline(yintercept = 0, color = "gray")
      p <- p + geom_text_repel(size = 4, max.overlaps = 50)
      p <- p + theme_minimal() 
      p <- p + theme(legend.position = "none",
                     axis.text = element_text(size = 18),
                     axis.title = element_text(size = 20)
                     )
      p <- p + labs(x = "First Dimension", y = "Third Dimension")
      d <- p + scale_color_manual(values = c('red', 'blue'))

      png(file = here("Plots", "ca-corr-plot.png"), width=1200, height=1200)
      (a + b) / c / d + 
         plot_annotation(tag_levels = 'a', tag_suffix = ')') & 
         theme(plot.tag = element_text(size = 26))
      dev.off()
```
